---
title: "Client-server Scope 3 workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Client-server Scope 3 workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Purpose

This vignette shows how the two-package architecture is used.

The architecture is:

```text
scope3supplier server package
  stores supplier uploads server-side
  stores emission factors server-side
  calculates emissions server-side
  exposes aggregate-only API endpoints

scope3federatedclient client package
  runs on another laptop
  knows only API URLs and supplier IDs
  receives aggregate outputs only
  produces the emissions summary/report
```

The client laptop does not need local supplier `.rds` files. It only needs the
client package and the API URL.

## 1. Install the Packages

During local development from this repository, install both packages from their
folders:

```{r install-local}
install.packages(
  "packages/scope3supplier",
  repos = NULL,
  type = "source"
)

install.packages(
  "packages/scope3federatedclient",
  repos = NULL,
  type = "source"
)
```

After you publish the packages to separate GitHub repositories, the client laptop
only needs the client package:

```{r install-github-client}
remotes::install_github("YOUR-GITHUB-USER/scope3federatedclient")
```

The server machine needs the server package:

```{r install-github-server}
remotes::install_github("YOUR-GITHUB-USER/scope3supplier")
```

## 2. Start the Server

Run this on the server machine, not on the client laptop.

For this prototype, we create server-side example data from the synthetic Excel
workbook. In production, suppliers would upload their own files through the
upload endpoint instead.

```{r server-setup}
library(scope3supplier)

create_example_server_data(
  workbook_path = "data/synthetic_scope3_category3_raw_data.xlsx",
  data_dir = "server-data"
)
```

This writes server-owned data under:

```text
server-data/
  emission_factors/
  suppliers/
```

Then start the Plumber API:

```{r server-start}
library(scope3supplier)

start_supplier_api(
  data_dir = "server-data",
  host = "127.0.0.1",
  port = 8080,
  minimum_group_size = 10
)
```

Leave this R process running. The server exposes endpoints like:

```text
GET /health
GET /emission-factors/version
GET /suppliers/SUPPLIER_APAC_ENERGY/record-count
GET /suppliers/SUPPLIER_APAC_ENERGY/emissions/total
GET /suppliers/SUPPLIER_APAC_ENERGY/emissions/by-activity
GET /suppliers/SUPPLIER_APAC_ENERGY/emissions/sufficient-statistics
```

## 3. Optional: Upload Supplier Data

Instead of creating example data, a supplier can upload a CSV file to the server.
The upload endpoint stores the data server-side and returns metadata only.

```{bash upload-example}
curl --data-binary @supplier.csv \
  "http://127.0.0.1:8080/suppliers/SUPPLIER_001/upload?filename=supplier.csv"
```

The uploaded file must contain at least:

```text
supplier_id
country
activity_type
quantity
unit
```

## 4. Run the Client From Another Laptop

This is the part you would run on a separate laptop with only the client package
installed.

```{r client-analysis}
library(scope3federatedclient)

suppliers <- data.frame(
  supplier_id = c(
    "SUPPLIER_AMERICAS_ENERGY",
    "SUPPLIER_APAC_ENERGY",
    "SUPPLIER_EUROPE_UTILITIES"
  ),
  base_url = rep("http://127.0.0.1:8080", 3)
)

result <- run_federated_analysis(
  suppliers = suppliers,
  output_dir = "outputs"
)
```

The result object contains:

```{r inspect-result}
names(result)
```

The main emissions summary is:

```{r federated-summary}
result$federated
```

To see the activity-level totals:

```{r activity-totals}
result$federated$activity_totals
```

To see the formula used by the architecture:

```{r formula}
emission_calculation_formula()
```

## 5. Files Written by the Client

If `output_dir = "outputs"` is supplied, the client writes:

```text
outputs/federated_report.md
```

If a reference/centralised result is also supplied, it also writes:

```text
outputs/centralised_federated_comparison.csv
```

In the normal another-laptop setup, the client does not need a centralised
reference result. It can still obtain the federated emissions summary from the
API.

## 6. What Data the Client Receives

The client receives only aggregate fields such as:

```text
supplier_id
n
total_emissions_tco2e
activity_type
sum_emissions_tco2e
sum_squares_emissions_tco2e2
```

The client does not receive:

```text
record_id
site_id
site_name
reporting_month
quantity
invoice_reference
raw activity records
```

## 7. How the Summary Is Computed

The server calculates record emissions locally:

```text
total_emissions_tco2e =
  quantity *
  (upstream_wtt_factor_kgco2e_per_unit + td_loss_factor_kgco2e_per_unit) /
  1000
```

The client calculates global results from sufficient statistics:

```text
global_total = sum(supplier totals)
global_mean = sum_supplier_sum_emissions / sum_supplier_n
global_sample_variance =
  (sum_supplier_sum_squares - sum_supplier_sum_emissions^2 / sum_supplier_n) /
  (sum_supplier_n - 1)
```

That is why the client can see the emission summary without seeing raw supplier
records.
